DATA 621 - Business Analytics and Data Mining

### 2020-11-20 HW5/HW5-Draft.Rmd
Finish basic EDA

# Get the names of the variables and which ones are numeric
nmtrn <- names(DT)

# Drop INDEX and target variables from name list
nmtrn <- nmtrn[-(1:2)]
```

@@

There are `r ntrnobs` observations. All of these predictors are numeric,
although some `LabelAppeal` and `STARS` could be considered ordered factors. For
the purposes of this assignment, it makes sense to treat every predictor as
numeric.

@@

```{r missing}
missingPreds <- transpose(DT[, lapply(.SD, function(x) {sum(is.na(x))}),
                             .SDcols = -c('INDEX', 'TARGET')],
                          keep.names = 'Predictors')
setnames(missingPreds, 'V1', 'Missing')
missingPreds[, Percentage := Missing / ntrnobs * 100]
setorder(missingPreds, -Missing)
kable(missingPreds, digits = 2L, caption = 'Predictors with Missing Observations')
```

@@

>>>>

# Get the names of the predictor variables
nmtrn <- names(DT)[-(1:2)]
nmtrnINT <- c('AcidIndex', 'LabelAppeal', 'STARS')
nmtrnDUB <- setdiff(nmtrn, nmtrnINT)
```

@@

There are `r ntrnobs` observations. All of these predictors are numeric, 
although `LabelAppeal`, `AcidIndex` and `STARS` all appear to be ordinal factors
and not true numerics. For the purpose of this assignment, we can treat them
as integers.

@@

```{r missingVal}
missingPreds <- transpose(DT[, lapply(.SD, function(x) {sum(is.na(x))}),
                             .SDcols = nmtrn],
                          keep.names = 'Predictors')
setnames(missingPreds, 'V1', 'Missing')
missingPreds[, Percentage := Missing / ntrnobs * 100]
setorder(missingPreds, -Missing)
kable(missingPreds, digits = 2L, caption = 'Predictors with Missing Observations')
```

@@

## Summary Statistics and Graphs
As all of the data are numeric, we can investigate the distributions of the
predictors both tabularly and graphically.

The numeric predictor variables have the following summary statistics, ignoring
missing values:

```{r statsN}
# Isolate numeric only predictors
predictorDT <- DT[, .SD, .SDcols = nmtrn]

# Melt them from wide to long format
predictorDTM <- melt(predictorDT, variable.name = 'metric',
                     value.name = 'value',  variable.factor = FALSE,
                     id.vars = NULL)

# Calculate summary statistics
statsN <- predictorDTM[, .(Mean = mean(value, na.rm = TRUE),
                           SD = sd(value, na.rm = TRUE),
                           Min = min(value, na.rm = TRUE),
                           Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
                           Median = median(value, na.rm = TRUE),
                           Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
                           Max = max(value, na.rm = TRUE),
                           IQR = IQR(value, na.rm = TRUE)), keyby = 'metric']

# Print the table
kable(statsN, digits = 3L, align = 'r',
      caption = "Summary Statistitics for Numeric Variables")
```

A kernel-smoothed density plot of the distributions of the non-integer numeric
predictors is below, followed by a histogram of the integral-valued predictors.

```{r graphsD, fig.height=5, fig.width=7}
# Using epanechnikov kernel
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnDUB], aes(x = value)) +
  geom_density(kernel = 'epanechnikov') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Kernel-smoothed Density of Numeric Predictors')
```

The numeric predictors all look to be basically symmetrical, but non-Gaussian in
that there is a strong spike near the median, but the tails on either side are
thicker than would be for a normal distribution with that low of a standard
deviation.

```{r graphsH, fig.height=2, fig.width=7}
# Freedman-Diaconis rule for bin widths
FDbin <- function(x) {
  result <- 2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
  return(ifelse(result == 0, 0.5, result))
}
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnINT], aes(x = value)) +
  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Histogram of Integral Predictors')
```

Of these three, `LabelAppeal` seems to be the most "normal" of the lot; `STARS`
and `AcidIndex` look to be more Poisson in shape.

### Correlations
The corrgram below graphically represents the correlations between the numeric
predictor variables, when ignoring the missing variables.

```{r corrgram, fig.width=6, fig.height=4}
corrplot::corrplot(cor(DT[, ..nmtrn], use = 'complete.obs'),
         method = 'ellipse', type = 'lower', order = 'hclust',
         hclust.method = 'ward.D2')
```

There is very little correlation between the variables. The only pairs with some
level correlation are:

  * `STARS` being positively correlated with `LabelAppeal`
    * This is interesting. Could it be that wine connoisseurs are impacted by
    the visual appearance of the label and not just the flavor?
  * `AcidIndex` having some positive correlation with `FixedAcidity`
    * One may have suspected a higher correlation, to be frank.
  * `AcidIndex` having some *negative* correlation with `STARS`
    * Wine reviewers may not like too much acidity, it seems.

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r statsN}
```
```{r graphsD}
```
```{r graphsH}
```
```{r corrgram}
```
```{r dataClean}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r m2clean}
```
```{r m3clean}
```
```{r model1DummyFreq}
```
```{r model1RunFreq}
```
```{r model1DummySev}
```
```{r model1RunSev}
```
```{r model2RunFreq}
```
```{r model2RunSev}
```
```{r model3RunBaseLog}
```
```{r model3RunFreq}
```
```{r model3RunBaseLin}
```
```{r model3RunSev}
```
```{r setUpCompTables}
```
```{r validateM1F}
```
```{r validateM1S}
```
```{r validateM2F}
```
```{r validateM2S}
```
```{r validateM3F}
```
```{r validateM3S}
```
```{r freqComp}
```
```{r sevComp}
```
```{r badSev}
```
```{r evalProcessClean}
```
```{r evalPredictFreq}
```
```{r evalPredictSev}
```
